Clustering time series based on probability distributions across temporal granularities

### 2021-11-17 main.Rmd
added standardization of wpd

@Gupta2021-hakear provides a method for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities. (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category. The reason for subsetting granularities in this way, is that, clustering algorithms perform badly in the presence of nuisance variables. Granularities that do not have some difference between categories are likely to be nuisance variables. It should be noted that all of the time series in a collection may not have the same set of significant granularities.<!--This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:--> This is the approach for generating a subset ($S_c$) of significant granularities across a collection of time series:

@@

The shape and scale of the distribution of the measured variable (e.g. energy usage) affects distance calculations. Skewed distributions need to be symmetrized. Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage. (Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly. Statistical transformations are employed to bring all of them to the same range or normalize each observation. This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS). NQT is used in @Gupta2021-hakear prior sorting granularities. <!-- $wpd$, which is the foundation of wpd-based distances. XX wpd was not yet introduced-->

@@

tibble::tibble(scenario, designs, niter, mu, nT) %>% #rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
  knitr::kable(format = "latex",
             escape = FALSE,
             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")%>%
  kableExtra::collapse_rows(columns = 3:5)
```

>>>>

@Gupta2021-hakear provides a distance measure ($wpd$) for determining the significance of a cyclic granularity, and a ranking of multiple cyclic granularities. (This extends to harmonies, pairs of granularities that might interact with each other.) We define "significant" granularities as those with significant distributional differences across at least one category. The reason for subsetting granularities in this way, is that, clustering algorithms perform badly in the presence of nuisance variables. Granularities that do not have some difference between categories are likely to be nuisance variables. It should be noted that all of the time series in a collection may not have the same set of significant granularities.<!--This is the approach for managing this generate a subset ($S_c$) of significant granularities across a collection of time series:--> This is the approach for generating a subset ($S_c$) of significant granularities across a collection of time series:

@@

The shape and scale of the distribution of the measured variable (e.g. energy usage) affects distance calculations. Skewed distributions need to be symmetrized. Scales of individuals needs to be standardized, because clustering is to select similar patterns, not magnitude of usage. (Organizing individuals based on magnitude can be achieved simply by sorting on a statistic like the average value across time.) <!--Observations often have a somewhat skewed time series distribution and their ranges might vary greatly. Statistical transformations are employed to bring all of them to the same range or normalize each observation. This is important because we are not interested in trivial clusters that vary in magnitudes, but rather in uncovering comparable patterns of distributional differences between categories.--> For the JS-based approaches, two data transformation techniques are recommended, normal-quantile transform (NQT) and robust scaling (RS).<!-- NQT is used in @Gupta2021-hakear prior sorting granularities.-- $wpd$, which is the foundation of wpd-based distances. XX wpd was not yet introduced--> While @Gupta2021-hakear already uses NQT when computing $wpd$, it could be useful to standardize it for the selected set of significant granularities prior to computing the distances.

@@

tibble::tibble(scenario, designs, niter, mu, nT) %>% rlang::set_names("scenario", "designs", "R", "$\\mu$", "T") %>% 
  #tibble::tibble(scenario, " designs" = designs, "R" =niter,  "$\\mu$" = mu, "T"=nT) %>% 
  knitr::kable(format = "latex",
             escape = FALSE,
             caption = "The range of parameters used for the validation study, for the three different scenarios, number of simulations ($R$) for each design, differences between means ($\\mu$) across granularities and series lengths ($T$).")
#%>%  kableExtra::collapse_rows(columns = 3:5)
```
